---
title: "Item combinations"
output: html_notebook
---
```{r loadpackages}
require(tidyverse)
require(gtools) #for permutations
```
Used to generate item combinations for 2020 version of Treasure Hunt.
#Update 5th Jan 2021
Generating spreadsheets for version in Registered Report with different constructions in Rep and Novel.

------------------------------------------------------------------------
#IMPORTANT UPDATE
1st August 2020 : I have set.seed to make spreadsheets reproducible.
If we want to generate alternative sheets, we will need to change this.
------------------------------------------------------------------------
1st August 2020
There was error in converting numeric codes for grid positions to letter codes. This process has been simplified and is now fixed.

Updated some parameters; 27/7/20

Also corrected to ensure fixed item is not in array; and repeated item is repeated across sessions

Tried with 3 repeated items rather than 4, so can get in more repetitions per session. However, this messes up counterbalancing: e.g. with 'above and below' you want half the repeated items to be above and half below. So reverted to 4.

so sequence is A N1 B N2 C N3 D N4 A N5 B N6 C N7 D N8 etc. This means if we have 3 rounds of A/B/C/D per session, we have 3 x 8  =24 items per session. 
This seems the maximum that we can expect children to do and so we have powered the study accordingly.

Need enough items for 4 sessions, i.e. 24 x 4 items altogether.  
The repeated item, by definition, is taken from a repeated set of 4

------------------------------------------------------------------------
## Difficulty level: this is now not varied.
 Everyone gets above/below or between
Two sets counterbalanced

Levels now are; 
#Table 1 (as per registered report)
Level 1	Level 2	Level 3	Level 4  

“The hen is between the star and the flag”	  
“The bed is above the tree and next to the jug"  

IWe create 2 frames x 4 (blocks) x 12 3-item combinations
This is 96 combinations.
Currently our stimulus pool is 32 pictured target words

Grid positions denoted in RowColumn format, ie..
11 12 13 14
21 22 23 24
31 32 33 34
41 42 43 44

Column then converted to letter
A1 B1 C1 D1
A2 B2 C2 D2
A3 B3 C3 D3
A4 B4 C4 D4

------------------------------------------------------------------------
From earlier versions:
Some questions that arise re item types.
a) With 'between' - we could include both vertical and horizontal grid positions - currently it is just horizontal
b) I have reworded level 4 items so it is not 2 sentences, but instead uses the format, e.g. the X is above the Y and next to the Z. This differs from Nicole's level 5.
c) Note that with the Repeated Items, it is currently set up to select at random the picture that will be pre-placed in the grid (where there is a preplaced item).  So the sentence is repeated but the display may vary. It would be possible to alter this - probably easiest to just do this at the final stage after spreadsheet created, at which point one could identify repeated items and just overwrite them so that each of the items was constant. But I think we are mostly interested in the impact of improved comprehension from hearing repeated sentence, and that to have this to some extent dissociated from the visual display is a good thing. This has now been clarified when describing methods

## Updates 3rd June 2020
Had a couple of minor edits:
1) spreadsheets now write to a 'sheets' folder- make things a little neater.
2) moved the creation of myspreadsheet to 337. It was previously specified later on preventing gridstart from running.
3) we add in sound generation for all spreadsheets.
4) in line 38, bed was coded as "bed " which meant spreadsheet wouldn't run in Gorilla.sc. Corrected.

```{r items}
set.seed(20) #make reproducible
myitems <- c( "clock","flag","car","cup","watch","kite","van","jug","wheel", "hen", "heart",  "lamp", "bat", "chain",  "spoon", "bed", "scarf", "pen", "leaf", "snail", "fly", "boat",  "crab", "drum", "box", "eye",  "leg", "fox",  "shirt", "bell",  "door", "plug",  "tree") 
itemcount<-length(myitems)
#NB These have been ordered so that 4 potentially confusible pairs occur at the front of the list; these are
# clock-watch
# flag-kite
# car-van
# cup-jug
# This ordering will make it relatively easy to ensure they don't co-occur in an item - see below

#get all permutations of 3 different items
permlist <- as.data.frame(permutations(n=32,r=3,v=1:itemcount,repeats.allowed=F))
```

This generates all possible 3-item permutations (ie ordered combinations) of the 32 words. There are 29760 of them, so we have plenty to choose from.

To create our experimental list, we could aim to select in quasirandom fashion, as follows:

a) Avoid any pictures that are phonologically/visually/semantically confusible occurring together in one sentence
b) Avoid adjacent sentences with any of the same words

To do (a) we'd need to specify lists of items that should not occur together, that we can check against

To do (b), just need a simple check.

So approach could be select 96 triplets by sampling at random, and accept each one if it meets criteria.

```{r sampletrios}
nrepstim <- 4 #these are N items that will be repeated per session
nreps <-3  #N repetitions of each repeated item within a session

itemtype<-c('R01','R02','R03','R04','N01','N02','N03','N04','N05','N06','N07','N08','N09','N10','N11','N12')

nlevels<-2 #just two levels which will be counterbalanced
nsess<-4
ntrio<-length(itemtype)#make data frame to hold trios
mytrio <- data.frame(matrix(NA,nrow=nlevels*nsess*ntrio,ncol=13))
#Note we'll create a frame to hold 5 pictures, but for now are just selecting 3. The other columns will hold distractors and are allocated later. Item 5 can be used later on to substitute in the choice array for the picture fixed in the grid.
colnames(mytrio)<-c('num1','num2','num3','num4','num5','pic1','pic2','pic3','pic4','pic5','itemtype','level','session')
myn <- nrow(permlist)
counter <- 0
nitems <- 12 #items per session per level (ie sentence type)
#NB we will use this later on when compiling Gorilla sheet. Note that because of repeated items, the N trios is less than this.
#Currently this is N items per session, with 4 repeated items presented 3 times plus one novel item for each
for(level in 1:nlevels){
  for(session in 1:nsess){
    for (item in 1:ntrio){
      counter <- counter+1
      #While loop here - these are dangerous: can get stuck in them!
      #But we start with mymatch set to 1 to ensure we enter the loop, and we only exit it when mymatch changes to zero - which it will do when there is no match with the prior 4 items
      mymatch<-1
      while(mymatch==1){
        thisone <- sample(myn,1) #we will return here to resample if there is overlap in the pictures within last 4 items
        mytrio[counter,1:3]<-permlist[thisone,1:3]
        #we'll check the prior 4 items to ensure no overlap in pics
        testlist<-vector()
        for (j in 1:3){
          myindex<-counter-j
          if(myindex>0) {
            testlist<-c(testlist,mytrio[myindex,1:3])
          }
        }
        w<-which(mytrio[counter,1:3]%in% testlist)
        
        #We now check if the items contain the pairing 1-5,2-6,3-7 or 4-8
        #These are pictures that may be confused, so to be avoided within an item
        w2=0 #initialise to flag no conflicting items
        w1<-which(mytrio[counter,1:3]<5) #any item code less than 5?
        if(length(w1)>0){
          badseq<-mytrio[counter,w1]+4 #items that conflict
          w2<-which(mytrio[counter,1:3]%in% badseq) #w2 will be a number from 1 to 3 if there are conflicting items
          if (length(w2)==0) {w2 <- 0} 
          #when length(w2) is > 0 we have 2 items in conflicct so need to resample
        }
        
        if((length(w)==0)&(w2==0)) #no overlap with prior items and no conflicting items!!
        {mymatch <-0} #can exit while loop because no overlap
      }
      
      mytrio$itemtype[counter]<-itemtype[item]
      mytrio$level[counter]<-level
      mytrio$session[counter]<-session
      
    }
  }
}
#For now we will ignore frames 1 and 2, so level will translate to 3 and 4
mytrio$level <- mytrio$level + 2
```

This seemed to work!  
We have 128 trios, with no overlap in items between items within 4-items apart. I think we also have avoided having conflicting pairs of picturs together in an item.

We need to also specify distractors for pic4 (and also pic5 - 5 pics needed if one is already placed! - this added 27/7/20) . 
We just add a picture that can be anything, provided it isn't already in the array.
Except they need also to be screened to avoid the confusible pairs.
This is done by just creating a list of 'used numbers' and as well as including those already selected for pic1-3, we include the confusible paired item in this list where necessary
This is also set up to avoid reusing distractors from previous trial - but NB actual presentation order is interspersed novel/random, so this is not going to work! Have left the code in, just in case we wanted to adjust at a later point. The main reason for adding distractors here is that we still have numeric codes for the stimuli. Script now modified so these are retained.

We need to use the same stimuli for repeated items across all sessions.
Easiest to do this by overwriting items.
We want to use session 1 versions for repeated items for all subsequent sessions.
I've moved this BEFORE the selection of distractors so that we get some variation in distractors for the repeated items. 

```{r copyrepeated}
myblock<-mytrio[1:nrepstim,1:nreps] #block for repeated sentences: to copy for later sessions
#NB this is hard coded and will need to change if we alter the N repeated sentences!
#Currently it is 12 because we have 4 versions of 3 repeated sentences
startrows<-which(mytrio$itemtype=='R01') #first repeated itemtype 

for (overwrite in startrows){
  mytrio[overwrite[1]:(overwrite[1]+nrepstim-1),1:nreps]<-myblock
}

```


```{r add.distractors}
for (p in 1:2){ #need to do this twice, so have 5 pics in all!
  for (i in 1:nrow(mytrio)){
    allnumbers <- 1:length(myitems)
    usednumbers <- c(mytrio$num1[i],mytrio$num2[i],mytrio$num3[i])
    if(p==2){usednumbers <- c(usednumbers,mytrio$num4[i])}
    #now add to this list the paired item for any that are in range 1-4
    u1<-which(usednumbers<5)
    if(length(u1)>0) {usednumbers<-c(usednumbers,usednumbers[u1]+4)}
    #now add to this list the paired item for any that are in range 5-8
    
    u2 <- intersect(which(usednumbers<9),which(usednumbers>4))
    if(length(u2)>0) {usednumbers<-c(usednumbers,usednumbers[u2]-4)}
    possiblenumbers <- allnumbers[-usednumbers]
    
    #Use different items 4 than in previous trial
    if (i>1){usednumbers<-c(usednumbers,mytrio$num1[(i-1)],mytrio$num2[(i-1)],mytrio$num3[(i-1)],mytrio$num4[(i-1)])}
    distractors <- sample(possiblenumbers,1)
    ifelse(p==1, mytrio$num4[i]<-distractors[1],mytrio$num5[i]<-distractors[1])
    
  }
}
```



So now we need to create the sentences and the spreadsheets.
Total of 7 sentence types.
Specify with X, Y, Z as placeholders and then replace with words from mytrio.

For each sentence type, can specify all possible locations


```{r makelookuptable}
#convert number codes to picture names -make copy of mytrio before doing this

mytrio$pic1<-myitems[mytrio$num1]
mytrio$pic2<-myitems[mytrio$num2]
mytrio$pic3<-myitems[mytrio$num3]
mytrio$pic4<-myitems[mytrio$num4]
mytrio$pic5<-myitems[mytrio$num5]
#We have 7 sentence frames at present; we will just allocate these numbers so we can use them in a lookup table, to create sentences in mytrio

#Note: this can get confusing - sentenceframe does not correspond directly to level. At levels 1 -2, we use frames 1 and 2; level 3-4 we use frame 3; level 4 we use frames 4-7.
sentenceframe<- c('The_X_is_above_the_Y',
                  'The_X_is_below_the_Y',
                  'The_X_is_underneath_the_Y_and_next_to_the_Z',
                  'The_X_is_next_to_the_Y_and_underneath_the_Z',
                  'The_X_is_on_top_of_the_Y_and_beside_the_Z',
                  'The_X_is_beside_the_Y_and_on_top_of_the_Z'
)
sentenceframe<-as.data.frame(sentenceframe)   



#Make grid that assigns sentence depending on item type and level
#We'll use this as a lookup table to create sentence for each item in mytrio
myt<-data.frame(table(mytrio$itemtype,mytrio$level)) #construct data frame from table
colnames(myt)<-c('itemtype','level','N')
myt<-myt[,-3] #lose 3rd column



myt$sentence <-0 #initialise
#We have to hard code the allocation of sentence types to levels

myt$level<-as.numeric(myt$level)+2 #convert from 1-2 to 3-4
#levels 1 and 2 - just alternate above and below (sentences 1-2)
# w<-which(myt$level<3)
# seq1 <- seq(from=1,to = max(w),by =2)
# myt$sentence[seq1] <- 1
# seq2<-seq1+1
# myt$sentence[(seq2)] <- 2 #below items

#level 3 - all are 'between'
w<-which(myt$level == 3)
myseq<-seq(from = min(w),to=max(w),by=2)
myt$sentence[w]<-3

myseq2<-myseq+1
myt$sentence[myseq2]<-4

#level 4: we have 4 possible sentence frames - just sequence through these.
w<-which(myt$level==4)
myseq<-seq(from = min(w),to=max(w),by=2)
myt$sentence[myseq]<-5

myseq2<-myseq+1
myt$sentence[myseq2]<-6


```





We'll compile specific sentences in a loop - just find the itemtype and level, and then can look up the sentence skeleton in myt.
At this point we also add the positions in the Gorilla grid, selecting at random from possible positions for this level. Note that once we know the position of X, the position of Y and Z can be deduced if we use a numeric code, so that you add 10 for each row in the grid and add 1 for each column

```{r compilesentences}
mytrio$sentence<-NA #initialise
mytrio$Xpos <-NA
mytrio$Ypos <-NA
mytrio$Zpos <-NA
mytrio$Y1pos <-NA #not sure if needed, but used in cases where more than one possible correct array
mytrio$Z1pos <-NA

# Grid positions denoted in RowColumn format, ie..
# 11 12 13 14
# 21 22 23 24
# 31 32 33 34
# 41 42 43 44

posXA <- c(23) #possible positions for X in 2-word 'above'
posXB <- c(32)#possible positions for X in 2-word 'below'
posXbetween <- c(12,13,22,23,32,33,42,43) #possible positions for X in between
posX5 <- c(21,22,23,31,32,33) #possible positions for X for level 5


for ( i in 1:nrow(mytrio)){
  myitem<-mytrio$itemtype[i]
  mylevel<-mytrio$level[i]
  w<-intersect(which(myt$itemtype==myitem),which(myt$level==mylevel))
  sentencetype<-myt$sentence[w]
  
  
  # allocate grid positions, depending on the sentence type
  if (sentencetype==1){
    mytrio$Xpos[i] <- posXA
    mytrio$Ypos[i]<-mytrio$Xpos[i]+10 #this is position immediately above
  }
  
  if (sentencetype==2){
    mytrio$Xpos[i] <- posXB
    mytrio$Ypos[i]<-mytrio$Xpos[i]-10 #this is position immediately below
  }
  
  if (sentencetype==3){
    mytrio$Xpos[i] <- sample(posXbetween,1)
    mytrio$Ypos[i]<-mytrio$Xpos[i]-1 #this is position to the left
    mytrio$Zpos[i]<-mytrio$Xpos[i]+1 #this is position to the right
    mytrio$Y1pos[i] <- mytrio$Zpos[i] #alternative correct arrangement
    mytrio$Z1pos[i] <- mytrio$Ypos[i]
  }
  if (sentencetype==4){
    #'The_X_is_above_the_Y_and_next_to_the_Z'
    mytrio$Xpos[i] <- sample(posX5,1)
    mytrio$Ypos[i]<-mytrio$Xpos[i]+10 #this is position immediately below
    mytrio$Zpos[i]<-mytrio$Xpos[i]+1
    mytrio$Z1pos[i]<-mytrio$Xpos[i]-1
  }
  
  if (sentencetype==5){
    #'The_X_is_below_the_Y_and_next_to_the_Z'
    mytrio$Xpos[i] <- sample(posX5,1)
    mytrio$Ypos[i]<-mytrio$Xpos[i]-10 #this is position immediately above
    mytrio$Zpos[i]<-mytrio$Xpos[i]+1
    
  }
  
  if (sentencetype==6){
    #'The_X_is_next_to_the_Y_and_above_the_Z'
    mytrio$Xpos[i] <- sample(posX5,1)
    mytrio$Ypos[i]<-mytrio$Xpos[i]+1
    mytrio$Zpos[i]<-mytrio$Xpos[i]+10 #this is position immediately below
    
  }
  
  if (sentencetype==7){
    #'The_X_is_next_to_the_Y_and_below_the_Z'
    mytrio$Xpos[i] <- sample(posX5,1)
    mytrio$Ypos[i]<-mytrio$Xpos[i]+1 
    mytrio$Zpos[i]<-mytrio$Xpos[i]-10 #this is position immediately below
    
  }
  
  #now compile the sentence
  mysent <-sentenceframe[sentencetype,]
  #Use gsub to just insert the relevant picture word - substitute for X, Y, Z
  mysent1 <- gsub("X", mytrio$pic1[i], mysent)
  mysent1 <- gsub("Y",mytrio$pic2[i],mysent1)
  if(mylevel>2)
  {  mysent1 <- gsub("Z",mytrio$pic3[i],mysent1)}
  mytrio$sentence[i]<-mysent1
  
  
}
```
At this point, pictures X Y and Z are in order 1, 2, and 3. 
Need to add list of pictures in random order - easiest just to do this with new columns , called choice
```{r randompictures}

mytrio$choice1<-NA
mytrio$choice2<-NA
mytrio$choice3<-NA
mytrio$choice4<-NA
mytrio$choice5<-NA


collist <- which(colnames(mytrio) %in% c('pic1','pic2','pic3','pic4','pic5'))
for (i in 1:nrow(mytrio)){
  nucollist <- which(colnames(mytrio) %in% c('choice1','choice2','choice3','choice4','choice5'))
}
for (i in 1:nrow(mytrio)){
  collist <- sample(collist) #scramble order of columns
  mytrio[i,nucollist]<-mytrio[i,collist]
}

```
Next we need to put these items into a spreadsheet in the format needed by Gorilla.
I'm taking one of Nicole's spreadsheets as a template.


```{r makespreadsheet}
spreadsheetcols <- c("A1", "B1", "C1", "D1", "A2", "B2", "C2", "D2", "A3", "B3", "C3", "D3", "A4", "B4", "C4", "D4", "Choice1", "Choice2", "Choice3", "Choice4", "Choice5","Clue1", "Answer1Choice", "Answer1Cell", "Answer2Choice", "Answer2Cell", "Answer3Choice", "Answer3Cell", "Answer4Choice", "Answer4Cell")
#We'll add columns that denote item identifier, session, and level
spreadsheetcols <-c("Item","Session","Level",spreadsheetcols)


#find columns that hold information about which slots in grid are open

#find columns that hold positions for correct answers
posstart <- which(colnames(mytrio)=='Xpos') #columnsin my trios with pos codes

#Make list of matrix positions that we can map onto cols A1, etc
matrixpos <- c(11,12,13,14,21,22,23,24,31,32,33,34,41,42,43,44)
alphanames <- c("A1", "B1", "C1", "D1", "A2", "B2", "C2", "D2", "A3", "B3", "C3", "D3", "A4", "B4", "C4", "D4")
#Easiest if we just make array with 44 spaces and put codes in where needed
myalphacodes <-vector()
for(a in 1:length(matrixpos)){
  myalphacodes[matrixpos[a]]<-alphanames[a]
}

#now populate spreadsheet: alternate repeated items (1-4) and novel.
for(mylevel in 3:4){

itemorder <- c(1,21,2,22,3,23,4,24,1,25,2,26,3,27,4,28,1,29,2,30,3,31,4,32)
if (mylevel==4){
  itemorder <- c(17,5,18,6,19,7,20,8,17,9,18,10,19,11,20,12,17,13,18,14,19,15,20,16)
}
#,3,11,4,12,1,13,2,14,3,15,4,16)
#,1,17,2,18,3,19,4,20,1,21,2,22,3,23,4,24)



maxm <-c(2,2,3,3,3) #this specifies how many elements to be considered when adding grid details (see below)

maxsession <- 4

for (mysession in 1:4){
 
    
    myspreadsheet <- data.frame(matrix(NA,ncol=length(spreadsheetcols),nrow=nitems*2))
    colnames(myspreadsheet) <- spreadsheetcols
    gridstart <- which(colnames(myspreadsheet)=='A1')
    gridcols <- gridstart:gridstart+15
    maxrow <- nrow(myspreadsheet)
    #Make list of items for this session
    thisbatch <- filter(mytrio,session==mysession)
    
    myspreadsheet$Item <- thisbatch$itemtype[itemorder]
    myspreadsheet$Session <- thisbatch$session[itemorder]
    myspreadsheet$Level <- thisbatch$level[itemorder]
    myspreadsheet$Choice1 <- thisbatch$choice1[itemorder]
    myspreadsheet$Choice2 <- thisbatch$choice2[itemorder]
    myspreadsheet$Choice3 <- thisbatch$choice3[itemorder]
    myspreadsheet$Choice4 <- thisbatch$choice4[itemorder]
    myspreadsheet$Choice5 <- thisbatch$choice5[itemorder]
    myspreadsheet$Clue1 <- thisbatch$sentence[itemorder]
    myspreadsheet$Answer1Choice <- thisbatch$pic1[itemorder]
    myspreadsheet$Answer1Cell <- thisbatch$Xpos[itemorder]
    myspreadsheet$Answer2Choice <- thisbatch$pic2[itemorder]
    myspreadsheet$Answer2Cell <- thisbatch$Ypos[itemorder]
    if(maxm[mylevel]>2){
      myspreadsheet$Answer3Choice <- thisbatch$pic3[itemorder]
      myspreadsheet$Answer3Cell <- thisbatch$Zpos[itemorder]
    }
    if(maxm[mylevel]>4){
      myspreadsheet$Answer4Choice <- thisbatch$pic2[itemorder]
      myspreadsheet$Answer4Cell <- thisbatch$Y1pos[itemorder]
      myspreadsheet$Answer5Choice <- thisbatch$pic3[itemorder]
      myspreadsheet$Answer5Cell <- thisbatch$Z1pos[itemorder]
    }
    startcell <- which(colnames(myspreadsheet)=='Answer1Cell')#find starting col for answers
    choicestart <- which(colnames(myspreadsheet)=='Choice1') #find starting col for choices
    for (n in 1:nrow(myspreadsheet)){
      starter <- startcell-2 #initialise
      #Levels 1 and 3 have pre-populated grid with one item
      myitem <- 0 #initialise - no prepopulated cells
      if(mylevel!=2){
        myitem<-sample(1:2,1)#either X or Y item
        mypiccols <-c((startcell-1),(startcell+1)) #cols for X and Y pictures
      }
      if(mylevel>2){
        myitem<-myitem+1 #select Y or Z
        mypiccols <-c(mypiccols,startcell+3) #include col for Z item
      } #select Y or Z
      
      for (m in 1:maxm[mylevel]){ #X, Y, Z, Y1 and Z1
        
        starter<-starter+2 #cell columns are 2 apart; on 1st run through, go back to Answer1Cell
        w<-which(matrixpos==myspreadsheet[n,starter]) #find matrix position
        if(length(w)>0){
          myspreadsheet[n,(w+gridstart-1)] <- 'open' #col corresponding to position for this item
          if(myitem==m){ #item to place in grid - 1st or 2nd in sentence
            placeditem<-myspreadsheet[n,mypiccols[myitem]] 
            placepos <- w+gridstart-1
            myspreadsheet[n,placepos]<-placeditem #place item in grid
            #if level 1, place open for positions above and below
            if (mylevel==1){
              #Place an 'open' cell above, i.e. 4 slots along
              myspreadsheet[n,(placepos-4)]<-'open'
              #Place an 'open' cell below, i.e. 4 slots along in other direction
              myspreadsheet[n,(placepos+4)]<-'open'
            }
            
            
            #Now ensure placed item is not in choice array (pics 1-4)
            placedchoice <- which(myspreadsheet[n,choicestart:(choicestart+3)]==placeditem)
            if(length(placedchoice)>0){
              myspreadsheet[n,(choicestart+placedchoice-1)]<-myspreadsheet[n,(choicestart+4)]
              #we write Choice 5 over the Choice for the placed item, so choice array doesn't contain the placed item.     
            }
          }
        }
        
      }
      
    } #next n
#------------------------------------------------------------------------------------   
#All rows done for this spreadsheet.
#New generic chunk to detect fixed items and ensure they are not in the answer columns.
#------------------------------------------------------------------------------------   
   
    myA1 <- which(colnames(myspreadsheet)=='A1')
    myans1 <-which(colnames(myspreadsheet)=='Answer1Choice')
    for (n in 1:nrow(myspreadsheet)){
    myplaced<-unname(unlist(myspreadsheet[n,myA1:(myA1+15)]))
    myanswers <- unname(unlist(myspreadsheet[n,myans1:(myans1+6)]))
    myfound <-intersect(myplaced,myitems)  #find item from original list of pictures
    if(length(myfound)>0){ #if no placed items this condition not met
      w<-which(myanswers==myfound)
      myspreadsheet[n,(myans1+w-1):(myans1+w)]<-NA #both AnswerChoice and AnswerCell set to NA
         }
} #next n
    
    
#------------------------------------------------------------------------------------   #Convert numeric codes for grid cells to alphanumeric 
#------------------------------------------------------------------------------------   

    ans1cell <- which(colnames(myspreadsheet)=='Answer1Cell')
    for (a in c(ans1cell,(ans1cell+2),(ans1cell+4))){
      w<-which(myspreadsheet[,a]>0)
      if(length(w)>0){
        myspreadsheet[w,a]<-myalphacodes[myspreadsheet[w,a]]
      }
      w1<-which(myspreadsheet[,a]==0)
      myspreadsheet[w1,a]<-NA
    } #next a (ie next cell column)
    
    spreadname<-paste0('./sheets/SheetSession', mysession, 'Level', '_', mylevel,'Rep.csv') # changed to work with gorilla mechanic
    
    w<-which(colnames(myspreadsheet)=='Choice5')
    myspreadsheet<-myspreadsheet[,-w]
    #With write.table we can ensure that NAs are saved as blanks
    write.table(myspreadsheet,spreadname,   na = "",
                row.names = FALSE,
                col.names = TRUE,
                sep = ",")
    
    
} #next level
  
} #next session

```


Make a little spreadsheet with just N items per level for testing with Gorilla

```{r maketestsheet}
#Now set to save 6 at each level, but could alter this
#Examples of levels 1,2,3,4 are saved 
thislevel <- 0 
Nperlevel <- 6
for (mylevel in c(1,2,3,4)){
  thislevel<-thislevel+1
  myfilename <- paste0('./sheets/sheetSession1Level_',mylevel,'.csv')
  testbit <- read.csv(myfilename,stringsAsFactors=F)
  if(mylevel==1){testsheet<-testbit[1:(4*Nperlevel),]}
  if(mylevel>1){
    theserows<-((thislevel-1)*Nperlevel+1):(thislevel*Nperlevel)
    testsheet[theserows,]<-testbit[1:Nperlevel,]
  }
  
}
#With write.table we can ensure that NAs are saved as blanks
write.table(testsheet,'./sheets/testsheet.csv',   na = "",
            row.names = FALSE,
            col.names = FALSE,
            sep = ",")
```


## Sound files

Now that we have the stimulus, we need to create stimuli. This is done via google voice. 

Dorothy, I had originally pushed the .json to github but Google then google emailed to say that this might comprmise our credentials. So I've taken it down. I'll send over if you need it.

I was not able to install googleLanguageR via normal interface, but I did manage it via
remotes::install_github("ropensci/googleLanguageR") - however, this seemed unhappy with versoin of R that I have

```{r soundGenerator}
makesounds<-1 #just make this chunk optional as it crashes if you don't have json properly installed
if (makesounds==1){
  library("googleLanguageR")
  # need json file in directory to start this
  gl_auth("sage-instrument-243609-abad49e474f9.json")
  
  # now pull all csv names in
  myfiles = list.files(path="./sheets/", pattern="*.csv", full.names=TRUE)
  # read them
  dat_csv = plyr::ldply(myfiles, read_csv)
  
  # now create the frame of info that we need
  items <- data.frame(matrix(NA,nrow=length(dat_csv$Clue1),ncol=2))
  colnames(items) <- c("label", "stim")
  items$label <- dat_csv$Clue1
  items$stim <- str_replace_all(items$label, fixed("_"), " ")
  # now run gl_talk to get the sounds
  mysent<-vector()
  for (i in 1:length(items$stim)) {
    gl_talk(items$stim[i],
            gender = "FEMALE", languageCode = "en-GB", speakingRate= .8, audioEncoding = "MP3", 
            output = (paste0("./mp3/",items$label[i],".mp3")))
  }
  write.csv(items$label,'./mp3/items_created.csv')
}
```
